A Tutorial on Tailored Power Simulations for Experimental Study Designs with Generalized Linear Mixed Models

### 2023-09-12 docs/manuscript.Rmd
changed some terminology

<!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
Currently available software packages for power simulations <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.  --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022]. 
A number of tutorials have been published describing how to perform such power simulation for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021]. 
However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->). This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->) [@brownIntroductionLinearMixedEffects2021]. 
The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure. The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial. Assumptions about the random effects include the assumption of normality<!--AK: Normality of what?--> and the covariance structure among the random effects<!--AK: Unclear what this is-->. Additionally, assumptions about the error structure, such as independence or correlation, must be specified. Interpreting these assumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters. However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [@maxwellSampleSizePlanning2008].

<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power simulation with GLMMs. Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.

>>>>

<!-- what are the existing options for a priori power analysis for GLMMs / where are existing tutorials limited -->
Currently available software packages for power analysis <!-- FP: because you removed the sentence on the difference between analytic and simulation based power analysis, it is now not clear why from now on we always talk about power simulation; TK: sorry if I have accidentally cut this distinction when merging versions, i think it is super important to have the definition between the two here.  --> do not include GLMMs, making it necessary to build data simulations tailored specifically to the study design [@westfallStatisticalPowerOptimal2014; @murayamaSummarystatisticsbasedPowerAnalysis2022]. 
A number of tutorials have been published describing how to perform such simulation-based power analysis for multilevel models [@arendStatisticalPowerTwolevel2019; @lafitSelectionNumberParticipants2021; @zimmerSampleSizePlanning2022; @brysbaertPowerAnalysisEffect2018; @kumleEstimatingPowerGeneralized2021; @debruineUnderstandingMixedEffectsModels2021]. 
However, most of these tutorials only cover linear mixed models (LMMs) and focus on the most common designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->). This narrow focus provides little guidance for researchers faced with more complex study designs (e.g., INSERT EXAMPLE <!--FP can you think of a good example?-->) [@brownIntroductionLinearMixedEffects2021]. 
The necessary assumptions for power simulations in GLMMs include assumptions about the distributional form of the outcome variable, the random effects, and the error structure. The distributional assumption specifies the family of distributions for the outcome variable, such as Gaussian, Poisson, or binomial. Assumptions about the random effects include the assumption of normality<!--AK: Normality of what?--> and the covariance structure among the random effects<!--AK: Unclear what this is-->. Additionally, assumptions about the error structure, such as independence or correlation, must be specified. Interpreting these assumptions entails understanding the underlying presumptions of the model and ensuring they align with the characteristics of the data being analyzed.
Existing tutorials often rely on heuristics for specifying variance components (e.g., the standard deviation of random intercepts) or assume that results from meta-analyses or data from pilot studies are available to determine plausible values for all model parameters. However, in practice, knowledge about those parameters from prior studies is often limited, specifying assumptions <!--AK: Again "assumptions".--> a practical challenge [@maxwellSampleSizePlanning2008].

<!-- Summary / Wrap-Up, pick up main arguments and what the paper provides, keep this short! -->
Based on the need for well-powered experimental research using GLMMs and the lack of tools to conduct corresponding power analyses, in this tutorial paper, we present a case study that serves as a practical demonstration of how to perform a simulation-based a priori power analysis with GLMMs. Thereby, we aim to equip researchers with the tools needed to simulate their own data and determine appropriate sample sizes for their own research.
